Untwisting the Boerdijk--Coxeter

### 2017-04-12 doc/TetrahelixUntwist.tex
proofread, adding V_general

\title{Untwisting the Tetrahelix}
\author{Robert L. Read \email{read.robert@gmail.com}
}

@@

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra.  Considering the edges of
  these tetrahedra as structural members, the resulting structure is attractive and
  inherently rigid, and therefore interesting to architects,
  mechanical engineers, and robotocists.  A formula is developed that matches the
  visually apparent helices forming the outer rails of the BC helix.
  This formula is generalized to a formula convenient to designers.
  Formulae for 
  computing the
  parameters that give edge-length minimax-optimal tetrahelices
  are given, defining a continuum of tetrahelices of varying curvature.
  The endpoints of the optimality of this continuum are the BC helix and
  a structure of zero curvature, the \emph{equitetrabeam}.
  Numerically finding the rail angle from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed. 
  An interactive tool for such design and experimentation is provided: \url{https://pubinv.github.io/tetrahelix/}.
  A formula for the inradius of optimal tetrahelices is given.
  Utility for static and variable geometry
  truss/space frame design and robotics is discussed.
\end{abstract}

@@

The vertices of the tetrahedra lie upon
three helices about the central axis.
The glussbot\cite{readglussbot} (or Tetrobot)\cite{TetrobotBook} uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
such as the three-d printable Song-Kwon-Kim\cite{song2003spherical} joint, or the CMS joint\cite{HamlinSandersonCMS}
which allow many members to come to a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

Imagining \cref{fig:closeup} as a static mechanical structure,
we observe that it is useful to the mechanical engineer or
robotocist because the structure remains an inherently rigid,
omni-triangulated space frame, which is
somewhat mechanically strong.
Imagine further in \cref{fig:closeup}, that each static edge was replaced with an
actuator that could dynamically become shorter or longer in response to electronic control,
and the vertices were a joint that supported sufficient angular displacement
for this to be possible. Such a machine is a Tetrobot\cite{TetrobotBook}.

@@

R.W. Gray's site\cite{graytetrahelix}, repeating a formula by Coxeter\cite{coxeter1985simplicial} in more accessible form, gives the Cartesian coordinates $\begin{bmatrix}
           x \\
           y \\
           z
         \end{bmatrix}$
for a counter-clockwise BC Helix:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos{n \theta_{bc}} $\\
   $ r_{bc} \sin{n \theta_{bc}} $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta_{bc} = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

@@

Such a helix can be written:
\begin{equation}
  \label{eq:colored}
H_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})}  $\\
   $ r_{bc} \sin{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})} $\\
   $ 3 h_{bc} (n + c/3)  $
  \end{tabular}
  \right ],
\text{where}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} \approx 0.5196 $\\
 $ h_{bc} = 1/\sqrt{10} \approx 0.3162 $ \\
 $ \theta_{bc} = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}

@@

Note in \cref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In \eqref{graycoxeter} and \eqref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$.
In this paper we assume the length of a rail
is always $1$ as a simplification. (We make the rail length a parameter in our JavaScript code in \texttt{tetrahelixmath.js} \cite{readtetrahelix}.)

The $H_{BCcolored}(n,c)$ formulation can be further clarified by rewriting directly in terms of the rail angle $\rho{bc}$ rather than $\theta_{bc}$.
Intuitively we seek an expression where $c/3$ is multiplied by a $1/3$ rotation plus the rail angle $\rho$.
We expand 
the expressions $\theta_{bc}$ and $\rho_{bc}$ in \eqref{eq:colored} and seek to isolate the term $c2\pi/3 $.
\begin{align*}
 c \theta_{bc}  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (c/3)  (3 \theta_{bc})  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
 (c/ 3) ((3 \theta_{bc} - 2 \pi)  + 2 \pi) &= \text{\{definition of $\rho_{bc}$\}...} \\  
  (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{algebra...} \\  
c  ( \rho_{bc} +  2 \pi) /3  \\
\end{align*}

@@

@@

We use the term \emph{tetrahelix} to mean any structure made of
vertices and edges which is isomorphic to the BC helix and in which the
vertices lie on three helices. We further demand that all edge lengths
be finite, as we are only interested in physically constructable tetrahelices.
By isomporphic we mean there is a one-to-one mapping between both
vertices and edges in the two tetrahelices.
One could consider various definitions of optimality for a
tetrahelix, but the most useful to us as robotocists extending the Tetrobot concept is to minimize the
maximum difference between any two edges, because the Tetrobot uses linear acutators of limit range of extension.

@@

Similarly, if the axes are not parallel, there is an edge of
unbounded length in the structure, so we exclude such a strcturue.

@@

 We show that in fact the nodes must be distributed in even thirds
 along the $z$-axis.

@@

 In principle in any 3 helices with the same axis of the same radius
 having any relative
 displacement along the $z$ axis there are 9 distinct edge classes.
 If when projecting all vertices on the the $z$-axis, the interval
 defined by the $z$ axis value of its endpoints contains no other vertices,
 we call it a \emph{one-hop} edge, and if it does contain another vertex we
 call it a \emph{two-hop} edge.
 Then there are 
 3 rail edges, 3 one-hop lengths between each pair of 3 rails, and 3 two hop
 lengths between each pair of three rails, where the two-hop length is at least
 the one-hop length.
 However, we have already shown the rail
 lengths are equal in any optimal tetrahelix.

@@

\Cref{equitetrabeam} shows the equitetrabeam, which is defined in \cref{sec:equitetrabeam},
but illustrates the one-hop and two-hop edge defintions.
The green edges are the two-hop edges and the purple edges are the one-hop edges. Note that the green
edges are slightly longer than the purple edges.

@@

\begin{proof}
    Consider the tetrahelix in which the vertices are evenly spaced at
    $h/3$ intervals on the $z$ axis. Every edge is either a rail edge,
    or it makes one hop, or it makes two hops. All of the one-hop
    edges are equal length.  All of the two-hop edges are equal
    length.

@@

    Any attempt to move any rail in either $z$ direction lengthens one two-hop edge to $t'$, where $t' > t$
    edge and shortens one one-hop edge $o' < o$. Let $B = \{o',t' \} \cup A$ be new edges.
    The minimax of $B$ is greater than the minimax of $A$ since there is a single rail length which cannot be both greater
    than $t'$ and $o'$ and less than $t'$ and $o'$.
    Therefore, any optimal tetrahelix has all one-hop edges between all rails equal to each other, and
    all two-hop edges equal to each other, and the $z$ distances between rails equal, and therefore
    $d/3$ from each other.
\end{proof}

@@

By \cref{eventhirds} every optimal tetrahelix has vertices lying on helices expressible in the form:
\[
V_{optimal}(n,c) =
\left [
  \begin{tabular}{c}
   $ r \cos(n \alpha +  c 2 \pi /3)$\\
   $ r \sin(n \alpha +  c 2 \pi /3) $\\
   $ \frac{d(n +c / 3)}{3}   $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $c \in \{0,1,2\}$
  \end{tabular}      
\]
where we have not yet investigated in the general case the relationships beteween $\alpha$, $r$, and $d$ in this formulation.
However, we understand that when $\alpha = 0$, the helices are degenerate, having curvature of $0$, and
we have the equitetrabeam.

@@

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to design the tetrahelix conveniently.
Please refer back to \cref{railanglefig}.
The pitch of the helix is an obvious choice, but is not defined when the
curvature is $0$, an important special case. The radius or the axial
distance between two nodes on the same rail are possible choices, but
perhaps the clearest choice is to build formulae that takes as thier
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle H(0,0) O H(0,1)$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of the tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality of counter-clockwise,
and $\chi = -1$ for clockwise.

 These quantities are related by the expression:

@@

Checking the important special case of the BC helix, we find that this equation
indeed holds true (treating $d$ in this equation as $3 h_{bc}$ as defined by
Gray and Coxeter, that is, $d_{bc} = 3h_{bc}$, where they are using it for the axial height from one node to
the next of a different color, but we use it to mean distance for the same color.

@@

Note that when $\rho = 0$ then $h_{\rho} = 1$, but $r_{\rho}$ is not determined by
\cref{railangle}.

@@

@@

If we look down the axis of an optimal tetrahelix as shown in \cref{axialview}, it happens that only one of the one-hop edges
(rendered in purpple in our software)
comes closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
A formula for the inradious of the tetrahelix is useful if you are designing it as a structure that bears something internally,
such as a firehose, a pipe, or
a ladder for a human. The inradius $r_{in}(\rho)$ of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  r_{in}(\rho) = r \sin{\frac{\pi - \rho}{6}}
\end{equation}
Which can be seen from the trigonometry of a diagram of the projected one-hop edges
connecting four sequentially numbered vertices:

@@

In \cref{fig:series}, the furthest tetrahelix is the optimal equitetrabeam.
Note that folding the outer edges of the paper upward or downward effectively
chooses the chirality of the equitetrabeam.

To the extent that we value tetrabeams (that is, tetrahelicies with a rail angle of $0$,
and therefore zero curvature) as mathematical or engineering objects,
we have motivated the development of $H_{general}$ as a transformation of $V(n)$ defined by
Equation \eqref{graycoxeter} from Gray and Coxeter. It is difficult to see how
the $V(n)$ 
formulation could ever give rise to a continuum producing the tetrabeam,
since setting the angle in that equation to zero can produce only collinear points.

@@

We observe that Equations \eqref{eqrhoopt} and \eqref{dopt} compute $r_{opt}$ and $d_{opt}$ which
create an optimal tetrahelix for any value rail angle $\rho$ between $0$, which
gives the equitetrabeam and
$\rho_{bc} \approx 35.43 \degree$, which gives the BC helix.

@@

The pitch of any tetrahelix 
where $\rho \neq 0$ is:
\begin{equation}
  \label{pitcheqn}
p(\rho) = \frac{2 \pi  \cdot d}{\rho}
\end{equation}
For a fixed $z$-axis travel $d$, this is trivial.
However, if one is computing 
$z$-axis travel from \eqref{dopt} the pitch is not simple.
It increases monotonically and smoothly with decreasing $\rho$, so
Equation \eqref{pitcheqn} can be easily solved numerically with a Newton-Raphson
solver, as we do on our website.
For a pitch at least $ p \geq \frac{3  \sqrt{2}  \pi}{\sqrt{5}\rho_{bc}} \approx 9.64 $,
using \eqref{dopt} produces minimax optimal tetrahelices.

@@

Perhaps surprisingly, the optimal untwisting is accomplished only by
changing the length of the two-hop member, leaving the one-hop member
and rail length equivalent within this continuum.\footnote{Before deriving Equation \eqref{eqrhoopt}, we created a continuum by
using a linear interpolation between the optimal radius for the
Equitetrabeam and the BC Helix. This minimax optimum of this simpler
approach was at most 1\% worse than the optimum computed by
\eqref{eqrhoopt}.}
 However, it should
be noted that an engineer or architect may also use $H_{general}$
directly and interactively, and that minimax length optimality is a
mathematical starting point rather than the final word on the beauty and utility of
physical structures. For example, a structural engineer might increase
radius in order to resist buckling. If an equitetrabeam were actually
used as a beam, an engineer might start with the optimal tetrabeam and
dilate it in one dimension to ``deepen'' the beam. Similarly, simple
changes curve the equitetrabeam into an ``arch''.

@@

\begin{figure}[H] %float with two figures
  \centering
     \includegraphics[width=0.9\textwidth]{figures/GlussBotBC.jpg}
     \caption{Glussbot in relaxed, or BC helix configuration}
     \includegraphics[width=0.9\textwidth]{figures/GlussBotEquitetrabeamCropped.jpg}
     \caption{The Equitetrabream: Fully Untwisted Glussbot}
\end{figure}

@@

The BC Helix is the end point of a continuum of tetrahelices, the
other end point being an uncurved tetrahelix with equilateral cross
section, constructed by changing the length of only those members
crossing the outside rails after hopping over the nearest
vertex. Under the condition of miminum maximum length difference of
all members in the system, all such tetrahelices have vertices evenly
spaced along the axis generated by a simple equation.  A mechanical
machine, such as a robot or a variable-geometry truss, that can change
the length of its members, can thus twist and untwist itself by
changing the length of the appropriate members to achieve any point in
the continuum optimally. With a numeric solution, a design may choose
a rotation angle and member lengths to obtain any desired pitch.

>>>>

\title{Untwisting the Boerdijk--Coxeter}
\author{Robert L. Read \email{read.robert@gmail.com}
}

@@

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra forming a helical column.  Considering the edges of
  these tetrahedra as structural members, the resulting structure is attractive and
  inherently rigid, and therefore interesting to architects,
  mechanical engineers, and robotocists.  A formula is developed that matches the
  visually apparent helices forming the outer rails of the BC helix.
  This formula is generalized to a formula convenient to designers.
  Formulae for 
  computing the
  parameters that give edge-length minimax-optimal tetrahelices
  are given, defining a continuum of tetrahelices of varying torsion.
  The endpoints of the optimality of this continuum are the BC helix and
  a structure of zero torsion, the \emph{equitetrabeam}.
  Numerically finding the rail angle from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed. 
  An interactive tool for such design and experimentation is provided: \url{https://pubinv.github.io/tetrahelix/}.
  A formula for the inradius of optimal tetrahelices is given.
  Utility for static and variable geometry
  truss/space frame design and robotics is discussed.
\end{abstract}

@@

The vertices of the tetrahedra lie upon
three helices about the central axis.
The glussbot\cite{readglussbot} (or Tetrobot)\cite{TetrobotBook} uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot uses mechanical actuators which can change their length, connected by special joints,
such as the 3D printable Song-Kwon-Kim\cite{song2003spherical} joint, or the CMS joint\cite{HamlinSandersonCMS}
which allow many members to meet in a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

Imagining \cref{fig:closeup} as a static mechanical structure,
we observe that it is useful to the mechanical engineer or
robotocist because the structure remains an inherently rigid,
omni-triangulated space frame, which is
mechanically strong.
Imagine further in \cref{fig:closeup}, that each static edge was replaced with an
actuator that could dynamically become shorter or longer in response to electronic control,
and the vertices were a joint that supported sufficient angular displacement
for this to be possible. An example of such a machine is a glussbot, shown in \cref{fig:glussbot}.

@@

R.W. Gray's site\cite{graytetrahelix}, repeating a formula by Coxeter\cite{coxeter1985simplicial} in more accessible form, gives the Cartesian coordinates $\begin{bmatrix}
           x \\
           y \\
           z
         \end{bmatrix}$
for a counter-clockwise BC Helix:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos{n \theta_{bc}} $\\
   $ r_{bc} \sin{n \theta_{bc}} $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} \approx 0.5196 $\\
 $ h_{bc} = 1/\sqrt{10} \approx 0.3162 $ \\
 $ \theta_{bc} = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

@@

Such a helix can be written:
\begin{equation}
  \label{eq:colored}
H_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})}  $\\
   $ r_{bc} \sin{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})} $\\
   $ 3 h_{bc} (n + c/3)  $
  \end{tabular}
  \right ],
\text{where}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta_{bc} = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}

@@

Note in \cref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In \eqref{graycoxeter} and \eqref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$.
In this paper we assume the length of a rail
is always $1$ as a simplification. (We make the rail length a parameter in our JavaScript code in
\url{https://github.com/PubInv/tetrahelix/blob/master/js/tetrahelix_math.js} \cite{readtetrahelix}.)

The $H_{BCcolored}(n,c)$ formulation can be further clarified by rewriting directly in terms of the rail angle $\rho_{bc}$ rather than $\theta_{bc}$.
Intuitively we seek an expression where $c/3$ is multiplied by a $1/3$ rotation plus the rail angle $\rho$.
We expand 
the expressions $\theta_{bc}$ and $\rho_{bc}$ in \eqref{eq:colored} and seek to isolate the term $c2\pi/3 $.
\begin{align*}
 c \theta_{bc}  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (c/3)  (3 \theta_{bc})  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
 (c/ 3) ((3 \theta_{bc} - 2 \pi)  + 2 \pi) &= \text{\{definition of $\rho_{bc}$\}...} \\  
  (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{algebra...} \\  
c  ( \rho_{bc} +  2 \pi) /3  \\
\end{align*}

@@

The \emph{pitch} of any tetrahelix, defined as the length of a complete revolution
where $\rho \neq 0$ is:
\begin{equation}
  \label{pitcheqn}
p(\rho) = \frac{2 \pi  \cdot d}{\rho}
\end{equation}

@@

We use the term \emph{tetrahelix} to mean any structure made of
vertices and edges which is isomorphic to the BC helix and in which the
vertices lie on three helices. We further demand that all edge lengths
be finite, as we are only interested in physically constructable tetrahelices.
By isomporphic we mean there is a one-to-one mapping between both
vertices and edges in the two tetrahelices.
One could consider various definitions of optimality for a
tetrahelix, but the most useful to us as robotocists working with the Tetrobot
concept is to minimize the
maximum difference between any two edges, because the Tetrobot uses
mechanical linear acutators with
limited range of extension.

@@

Similarly, if the axes are not parallel, there is an edge of
unbounded length in the structure, so we do not consider such cases.

@@

 We show that in fact the nodes must be distributed in even thirds
 along the $z$-axis, as in fact they are in the regular BC helix.

@@

 In principle in any 3 helices with the same axis of the same radius
 having any relative
 displacement along the $z$ axis there are 9 distinct edge classes.
 If when projecting all vertices on the the $z$-axis, the interval
 defined by the $z$ axis value of its endpoints contains no other vertices,
 we call it a \emph{one-hop} edge, and if it does contain another vertex we
 call it a \emph{two-hop} edge.
 Then there are 
 3 rail edges, 3 one-hop lengths between each pair of 3 rails, and 3 two hop
 lengths between each pair of 3 rails, where the two-hop length is at least
 the one-hop length.
 However, we have already shown the rail
 lengths are equal in any optimal tetrahelix.

@@

\Cref{equitetrabeam} shows the equitetrabeam, which is defined in \cref{sec:equitetrabeam},
but also conveniently illustrates the one-hop and two-hop edge defintions.
The green edges are the two-hop edges and the purple edges are the one-hop edges. Note that the green
edges are slightly longer than the purple edges.

@@

\begin{proof}
    Consider the tetrahelix in which the vertices are evenly spaced at
    $d/3$ intervals on the $z$ axis. Every edge is either a rail edge,
    or it makes one hop, or it makes two hops. All of the one-hop
    edges are equal length.  All of the two-hop edges are equal
    length.

@@

    Any attempt to move any rail in either $z$ direction lengthens one two-hop edge to $t'$, where $t' > t$
    and shortens one one-hop edge $o' < o$. Let $B = \{o',t' \} \cup A$ be new edges.
    The minimax of $B$ is greater than the minimax of $A$ since there is a single rail length which cannot be both greater
    than $t'$ and $o'$ and less than $t'$ and $o'$.
    Therefore, any optimal tetrahelix has all one-hop edges between all rails equal to each other, and
    all two-hop edges equal to each other, and the $z$ distances between rails equal, and therefore
    $d/3$ from each other.
\end{proof}

@@

By \cref{eventhirds} every optimal tetrahelix has vertices lying on helices expressible in the form:
\[
V_{optimal}(n,c) =
\left [
  \begin{tabular}{c}
   $ r \cos(n \alpha +  c 2 \pi /3)$\\
   $ r \sin(n \alpha +  c 2 \pi /3) $\\
   $ \frac{d(n +c / 3)}{3}   $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $c \in \{0,1,2\}$
  \end{tabular}      
\]
where we have not yet investigated in the general case the relationships beteween $\alpha$, $r$, and $d$ in this formulation.
However, we understand that when $\alpha = 0$, the helices are degenerate, having torsion of $0$, and
we have the equitetrabeam.

@@

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to design the tetrahelix conveniently.
Please refer back to \cref{railanglefig}.
The pitch of the helix is an obvious choice, but is not defined when the
torsion is $0$, an important special case. The radius or the axial
distance between two nodes on the same rail are possible choices, but
perhaps the clearest choice is to build formulae that takes as their
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle H(0,0) O H(0,1)$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of the tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality of counter-clockwise,
and $\chi = -1$ for clockwise.

 The quantities $\rho,r,d$ are related by the expression:

@@

Checking the important special case of the BC helix, we find that this equation
indeed holds true (treating $d$ in this equation as $3 h_{bc}$ as defined by
Gray and Coxeter, that is, $d_{bc} = 3h_{bc}$, where they are using
$h$ for the axial height from one node to
the next of a different color, but we use $d$ to mean distance between nodes of
the  same color.

@@

Note that when $\rho = 0$ then $d_{\rho} = 1$, but $r_{\rho}$ is not determined by
\cref{railangle}.

@@

Writing as robotocists, the ``colored'' approach of thinking of three rails seems natural to us, in part because our application
is focused on the edges, rather than any sort of interpretaion of the tetrahedra as solids. \Cref{eq:general} is in terms of helices that are close
to what you see when you look at a tetrahelix constructed from edges.
Having proved \cref{generalformulaoptimal}, it becomes plain that one could also use
 a generalization of the orginal Coxeter/Gray formulation which uses a single helix to define all vertices, and a mathematician, chemist,
or solid modeler might prefer that. This formulation is equivalent to \Cref{eq:general}, but uses $\theta = \frac{\rho+ 2\pi}{3}$,
the rotation between rails, and triples the rotation rate. The same constraint on $r_{opt}$ and $d_{opt}$ holds if the user seeks minimax optimality.

\begin{align*}
\numberthis  \label{eq:coxetergraygeneral}  
V_{general}(\chi,k,\theta,d_{\theta},r_{\theta}) &= 
 \left [
  \begin{tabular}{c}
   $ r_{\theta} \cos(\chi k \theta) $\\
   $ r_{\theta}  \sin(\chi k \theta) $\\
   $ d_{\theta} k / 3 $
  \end{tabular}
  \right ] \\
\text{where: } &
\begin{tabular}{c}
  $   1 = d_{\theta}^2 + 4 r_{\theta}^2 (\sin{ \frac{3\theta - 2\pi}{2}})^2 $ \\
    $\chi \in \{-1,1\}$ \\  
\end{tabular}
\end{align*}

Thus for any choice or $d$ and $r$ : $V_{general}(\chi,k,\theta,d,r) = H_{general}(\chi,n,c,\rho,d,r)$ when
$\theta = \frac{\rho+ 2\pi}{3}$ and $k = 3n + c$,
and both produce a minimax optimal tetrahelix for appropriate choices of $r$ and $d$ for
$\frac{2\pi}{3} \leq \theta \leq \theta_{bc}$ or $0 \leq \rho \leq \rho_{bc}$.

@@

If we look down the axis of an optimal tetrahelix as shown in \cref{axialview}, it happens that only one of the one-hop edges
(rendered in purple in our software)
comes closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
A formula for the inradious of the tetrahelix is useful if you are designing it as a structure that bears something internally,
such as a firehose, a pipe, or
a ladder for a human. The inradius $r_{in}(\rho)$ of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  r_{in}(\rho) = r \sin{\frac{\pi - \rho}{6}}
\end{equation}
Which can be seen from the trigonometry of a diagram of the projected one-hop edges
connecting four sequentially numbered vertices:

@@

In \cref{fig:series}, the furthest tetrahelix is the optimal equitetrabeam.

To the extent that we value tetrabeams (that is, tetrahelices with a rail angle of $0$,
and therefore zero curvature and torsion) as mathematical or engineering objects,
we have motivated the development of $H_{general}$ as a transformation of $V(n)$ defined by
Equation \eqref{graycoxeter} from Gray and Coxeter. It is difficult to see how
the $V(n)$ 
formulation could ever give rise to a continuum producing the tetrabeam,
since setting the angle in that equation to zero can produce only collinear points.

@@

We observe that Equations \eqref{eqrhoopt} and \eqref{dopt} compute $r_{opt}$ and $d_{opt}$ which
create an optimal tetrahelix for any rail angle $\rho$ between $0$, which
gives the equitetrabeam and
$\rho_{bc} \approx 35.43 \degree$, which gives the BC helix.

@@

 The pitch of a helix (see \cref{pitcheqn}, for a fixed $z$-axis travel $d$,
 is trivial.
However, if one is computing 
$z$-axis travel from \eqref{dopt} the pitch is not simple.
It increases monotonically and smoothly with decreasing $\rho$, so
Equation \eqref{pitcheqn} can be easily solved numerically with a Newton-Raphson
solver, as we do on our website.
For a pitch at least $ p \geq \frac{3  \sqrt{2}  \pi}{\sqrt{5}\rho_{bc}} \approx 9.64 $,
using \eqref{dopt} produces minimax optimal tetrahelices.

@@

Perhaps surprisingly, the optimal untwisting is accomplished only by
changing the length of the two-hop member, leaving the one-hop member
and rail length equivalent within this continuum.\footnote{Before deriving Equation \eqref{eqrhoopt}, we created a continuum by
using a linear interpolation between the optimal radius for the
Equitetrabeam and the BC Helix. This minimax optimum of this simpler
approach was at most 1\% worse than the optimum computed by
\eqref{eqrhoopt}.}
 However, it should
be noted that an engineer or architect may also use $H_{general}$
directly and interactively, and that minimax length optimality is a
mathematical starting point rather than the final word on the beauty and utility of
physical structures. For example, a structural engineer might increase
radius past optimality in order to resist buckling.
%% Add a figure here of a deep beam and of an arch.

If an equitetrabeam were actually
used as a beam, an engineer might start with the optimal tetrabeam and
dilate it in one dimension to ``deepen'' the beam. Similarly, simple
length changes curve the equitetrabeam into an ``arch''.

@@

\begin{figure}[H] %float with two figures
  \centering
     \includegraphics[width=0.9\textwidth]{figures/GlussBotBC.jpg}
     \caption{Glussbot in relaxed, or BC helix configuration}
     \includegraphics[width=0.9\textwidth]{figures/GlussBotEquitetrabeamCropped.jpg}
     \caption{The Equitetrabream: Fully Untwisted Glussbot in Hexapod Configuration}
     \label{fig:glussbot}
\end{figure}

By untwisting the tetrahelix so that is has a planar surface resting on the ground,
we may consider each vertex touching the ground a foot or pseudopod. A robot can thus
become a hexapod or $n$-pod robot, and the already well-developed approaches to
hexapod gaits may be applied to make the robot walk or crawl.

@@

The BC Helix is the end point of a continuum of tetrahelices, the
other end point being an untwisted tetrahelix with equilateral cross
section, constructed by changing the length of only those members
crossing the outside rails after hopping over the nearest
vertex. Under the condition of miminum maximum length difference of
all members in the system, all such tetrahelices have vertices evenly
spaced along the axis generated by a simple equation.  A mechanical
machine, such as a robot or a variable-geometry truss, that can change
the length of its members, can thus twist and untwist itself by
changing the length of the appropriate members to achieve any point in
the continuum optimally. With a numeric solution, a design may choose
a rotation angle and member lengths to obtain any desired pitch.
